---
title: "02_dataset_creation_full"
author: "Miguel Armengol"
date: "2/20/2019"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environment

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(feather)
library(knitr)
library(plotly)
library(sqldf)
library(zoo)
library(data.table)
library(summarytools)
library(imputeTS)
library(pracma)
library(entropy)
```

# abp_mean_full_surgery

## Filtering out

### Duration Surgery (min) calculation 

#### First and End Times of VitalSigns, VitalSigns_Times creation

*Run 011_fixing_surgery_times.Rmd before going on (the 011 file won't be included on github since it contains PHI)*

```{r}
library(dplyr)

    abp_mean_First <- abp_mean %>% 
    group_by(Case_Name) %>% 
    dplyr::select(c(Case_Name,Time)) %>%
    arrange(Time)%>%
    filter(!duplicated(Case_Name)) 
    colnames(abp_mean_First) <- c('Case_Name', 'abp_mean_First_Time')
    
    abp_mean_Last <- abp_mean %>% 
    group_by(Case_Name) %>% 
    dplyr::select(c(Case_Name,Time)) %>%
    arrange(desc(Time))%>%
    filter(!duplicated(Case_Name)) 
    colnames(abp_mean_Last) <- c('Case_Name', 'abp_mean_Last_Time')
    
    abp_mean_Times<-merge(abp_mean_First, abp_mean_Last)
    
```

##### Duration times

```{r}
abp_mean_Times['duration_Surgery_min']<-as.numeric(difftime(abp_mean_Times$`abp_mean_Last_Time`,abp_mean_Times$`abp_mean_First_Time`), units="mins")
```

### Creating lag and diff columns 

```{r message=FALSE, warning=FALSE}
abp_mean<-abp_mean%>%
  group_by(Case_Name)%>%
   mutate(
      time_lag=lag(Time)
     ,diff= abs(as.numeric(time_lag-Time))
   )
```

### Ploting sampling time

```{r}
abp_mean%>%
group_by(Case_Name)%>%
filter(diff>5*60)%>%
  summarise(n_distinct(Case_Name))
``` 

The median surgery length is 283.5 minutes.
We have a total of 174 (3%) cases with missing sampling times > 5 minutes (big 'holes' without data).
*We have decided to exclude those cases.*

### Excluding cases with missing sampling gaps  

We are excluding gaps > 5 minutes or surgery length < 60 minutes.

According to our previous work with this data-set, times-tamps indicating surgeries lasted less than 60 minutes might be an error from the monitor that is recording the vitals, thus we are excluding those cases. 

```{r}
# Creating the not in function

'%!in%' <- function(x,y)!('%in%'(x,y))

cases_to_exclude<-union_all(
abp_mean%>%
group_by(Case_Name)%>%
filter(diff>5*60)%>%
  select(Case_Name)
,
abp_mean_Times%>%
filter(duration_Surgery_min <= 60)%>%
  select(Case_Name)
)



cases_to_exclude<-unique(cases_to_exclude$Case_Name)
print('Number of Initial cases')
a<-n_distinct(abp_mean$Case_Name)
a
abp_mean_filtered<-abp_mean%>%
  filter(Case_Name %!in% cases_to_exclude)
print('Number of Cases Excluded due to gaps > 5 minutes or surgery length < 60 minutes.')
b<-n_distinct(abp_mean_filtered$Case_Name)
a-b
print('Number of Final cases')
b

```

5979 - 186 (missing sampling gaps > 5 minutes or surgery length < 60 minutes) = 5793

### Expanding data points

Expanding data points when several missing time-stamps are found

```{r}
  abp_mean_full_times<-as.data.frame(matrix(0,ncol = 1, nrow = 0))
  colnames(abp_mean_full_times)<-c('Time')

  for (i in 1:nrow(abp_mean_Times)) {
  correct_sequence<-as.data.frame(seq(abp_mean_Times$abp_mean_First_Time[i],abp_mean_Times$abp_mean_Last_Time[i], by = '15 secs'))
  correct_sequence['Case_Name']<-abp_mean_Times$Case_Name[i]
  abp_mean_full_times<-rbind(correct_sequence,abp_mean_full_times)
  }
  colnames(abp_mean_full_times)[1]<-'Time'
  
# excluding cases as previously defined, also from this dataframe
  abp_mean_full_times<-abp_mean_full_times%>%
  filter(Case_Name %!in% cases_to_exclude)   

abp_mean_expanded<-left_join(abp_mean_full_times,abp_mean_filtered%>%select('Case_Name','Time','Value'), by = c('Case_Name','Time'))
```

### Excluding cases with invalid cpb, precpb, postcpb or outsidecpb data-points

```{r}

# cases with valid cpb data
abp_mean_valid_cases1<-inner_join(  abp_mean_expanded,  AllPeriodsDF)%>%
  filter(Time>=CPB_Begin_Time & Time<=CPB_End_Time)%>%
  distinct(Case_Name)

# cases with valid outsidecpb data
abp_mean_valid_cases2<-inner_join(  abp_mean_expanded,  AllPeriodsDF)%>%
  filter(Time< CPB_Begin_Time | Time>CPB_End_Time)%>%
  distinct(Case_Name)

# cases with valid precpb data
abp_mean_valid_cases3<-inner_join(  abp_mean_expanded,  AllPeriodsDF)%>%
  filter(Time< CPB_Begin_Time)%>%
  distinct(Case_Name)

# cases with valid postcpb data
abp_mean_valid_cases4<-inner_join(  abp_mean_expanded,  AllPeriodsDF)%>%
  filter(Time > CPB_Begin_Time)%>%
  distinct(Case_Name)

valid_cases<-inner_join(abp_mean_valid_cases1,abp_mean_valid_cases2)
valid_cases<-inner_join(valid_cases,abp_mean_valid_cases3)
valid_cases<-inner_join(valid_cases,abp_mean_valid_cases4)
```

### Selecting only valid cases from our dataset

```{r}
print('Number of Initial cases')
a<-n_distinct(abp_mean_expanded$Case_Name)
a


abp_mean_expanded<-inner_join(abp_mean_expanded , valid_cases)

print('Number of Cases Excluded due to Inadequate Perioperative Data')
b<-n_distinct(abp_mean_expanded$Case_Name)
a-b

print('Number of Final cases')
n_distinct(abp_mean_expanded$Case_Name)
```


### Artifact Removal

Remove artifacts using the following rules, in order: 

(a) SBP (Systolic Blood Pressure) greater than or equal to 300 or SBP less than or equal to 20 mmhg.
(b) SBP less than or equal to DBP + 5 mmhg.
(c) DBP (Diastolic Blood Pressure) less than or equal to 5 mmhg or DBP greater than or equal to 225 mmhg.

_Ref: Salmasi, V., Maheshwari, K., Yang, D., Mascha, E. J., Singh, A., Sessler, D. I., & Kurz, A. (2017). Relationship between Intraoperative Hypotension, Defined by Either Reduction from Baseline or Absolute Thresholds, and Acute Kidney and Myocardial Injury after Noncardiac SurgeryA Retrospective Cohort Analysis. Anesthesiology: The Journal of the American Society of Anesthesiologists, 126(1), 47-65._

```{r}
#MAP artifact removal
abp_mean_expanded$Value[abp_mean_expanded$Value <=  0]<-NA
abp_mean_expanded$Value[abp_mean_expanded$Value >=  255]<-NA
```


### Imputing intra-surgery missing data

#### Median imputation

We used median imputation since sometimes the first values of a dataframe are all artifacts (so they were removed), therefore using LPV would give NA to all of them.

```{r}
impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
abp_mean_expanded<-abp_mean_expanded%>%
  group_by(Case_Name)%>% 
    mutate(
        Value_imputed = as.integer(impute.median(Value))
    )
```

LVP and linear interpolation methods can also be tried for imputation.

### Reshaping

```{r}
# I'm creating a counter column so we know how many datapoints we have per case (we'll later use this for reshaping the dataframe)
abp_mean_expanded_count<-abp_mean_expanded %>%
  group_by(Case_Name) %>%
  mutate(
    ones = 1,
    counter = cumsum(ones)
    )

# Reshaping the table so we have 1 row per patient and one column each data point
abp_mean_reshaped<-dcast(abp_mean_expanded_count, Case_Name ~ counter, value.var = "Value_imputed")
```

Each observation (row) corresponds to one case, there are 3650 data points (variables=columns), 1 every 15 seconds. 
This is because the longest case contains 3650 data points using ST 15 seconds. 
Cases shorter than this one will contain NA from the last time-stamp of the surgery until data-point number 3650.

'data.frame':	5793 obs. of  3651 variables:
 $ Case_Name: int  1 2 3 4 ...
 $ 1        : int  66 63 23 216 197 64 72 73 78 39 ...
 $ 2        : int  66 63 121 69 93 64 72 73 78 98 ...
 $ 3        : int  66 63 193 69 91 64 72 73 78 125 ...
 $ 4        : int  66 63 103 67 91 64 72 73 78 91 ...
 $ 5        : int  66 63 103 70 90 64 72 73 78 79 ...
 $ 6        : int  66 63 103 73 90 64 72 73 78 77 ...
 $ 7        : int  66 63 101 70 87 64 5 73 73 81 ...
 $ 8        : int  66 63 108 69 83 64 85 73 70 82 ...
 $ 9        : int  66 63 108 67 86 64 100 73 72 85 ...
 $ 10       : int  66 63 106 67 85 64 101 73 75 85 ...
 $ 11       : int  66 63 111 68 84 64 104 73 76 81 ...
 $ 12       : int  66 63 122 69 90 56 100 73 78 82 ...
 $ 13       : int  66 63 129 78 99 60 97 73 76 81 ...
 $ 14       : int  115 63 112 78 97 78 96 73 78 80 ...
 $ 15       : int  86 63 99 70 93 77 94 198 76 82 ...

### Imputing data with a special value.

Instead of removing the rows with NaN values when surgery finishes, we can replace all NaN values with a specific value that does not appear naturally in the input, such as -1.
The problem can be modeled as-is and we can encourage the model to learn that a specific value means “missing.” 

Ref: https://machinelearningmastery.com/handle-missing-timesteps-sequence-prediction-problems-python/

#### Replacing NA with -1

```{r}
abp_mean_reshaped[is.na(abp_mean_reshaped)]<-(-1) 
```

### Entropy

We are not using entropy anymore.

```{r eval=FALSE, include=FALSE}
abp_mean_reshaped['entropy']<-round(apply(abp_mean_reshaped[2:ncol(abp_mean_reshaped)], 1 , entropy ) , 2)
```

189 (3%) of the cases are outliers according the the boxplot upper fence.

_In the end we are not removing data beyond outliers_

#### Removing cases beyond outliers

_In the end we are not removing data beyond outliers_

```{r eval=FALSE, include=FALSE}
abp_mean_no_outliers<-abp_mean_reshaped[,1:1791] # we need to add 1 (1790+1) since the first column is Case_Name
```

#### Imputing after surgery missing data

*We are not imputing after surgery missing data anymore*

Linear interpolation imputation works vertically, so we transpose the dataframe to have one column per case, instead of one row per case.

```{r eval=FALSE, include=FALSE}
abp_mean_no_outliers_t<-as.data.frame(t(abp_mean_no_outliers[,-1]))
colnames(abp_mean_no_outliers_t) <- abp_mean_no_outliers[,1]
abp_mean_no_outliers_t_imputed<-na.interpolation(abp_mean_no_outliers_t, option = "linear")

abp_mean_no_outliers_imputed<-as.data.frame(t(abp_mean_no_outliers_t_imputed))
abp_mean_no_outliers_imputed['Case_Name']<-rownames(abp_mean_no_outliers_imputed)
row.names(abp_mean_no_outliers_imputed)<-1:nrow(abp_mean_no_outliers_imputed)

# re-ordering columns, we want Case_Name to be the first column

desired_order<-1791
desired_order<-c(desired_order,1:1790)
abp_mean_no_outliers_imputed<-abp_mean_no_outliers_imputed[,desired_order]
```



### Joining with other data-sets
 
```{r}
abp_mean_all_data<-inner_join(abp_mean_reshaped,FinalHemodynamicsDataset%>%
                          dplyr::select( Case_Name
                                        ,Gender
                                        ,Age
                                        ,Category
                                        
                                        ,duration_Surgery_min
                                        ,duration_CPB_min
                                        ,duration_outsideCPB_min
                                        ,duration_preCPB_min
                                        ,duration_postCPB_min 
                                        
                                        ,STS_RiskAlgorithm_multiple_imp
                                        ,STS_RiskAlgorithm_median_imp
                                        ,STS_RiskAlgorithm
                                        ,total_pressors_formula_mg
                                        
                                        # categorical variables
                                        ,Q_total_pressors_formula_mg
                                        ,T_STS_RiskAlgorithm_median_imp
                                        ,T_STS_RiskAlgorithm_multiple_imp
                                        ,Q_AUC65MAP_total_mmHgmin
                                        ,Q_TWA65MAP_total_mmHg
                                        ,Q_AUC65MAP_CPB_mmHgmin
                                        ,Q_AUC65MAP_outsideCPB_mmHgmin
                                        ,Q_TWA65MAP_CPB_mmHg
                                        ,Q_TWA65MAP_outsideCPB_mmHg
                                        
                                                                                
                                        # TWA variables
                                        ,TWA65MAP_total_mmHg
                                        ,TWA65MAP_CPB_mmHg
                                        ,TWA65MAP_outsideCPB_mmHg
                                        ,TWA65MAP_preCPB_mmHg 
                                        ,TWA65MAP_postCPB_mmHg 
                                        

                                                                                                                                                                ,delta_HCT_pct_median_imp
                                        ,mean_ef_median_imp
                                        ,Cross_Clamp_Time_imp
                                        
                                        ,AUC65MAP_total_mmHgmin
                                        ,AUC65MAP_CPB_mmHgmin
                                        ,AUC65MAP_outsideCPB_mmHgmin
                                        ,AUC65MAP_preCPB_mmHgmin 
                                        ,AUC65MAP_postCPB_mmHgmin                                        
                                        # Outcomes
                                        ,Any_MAE
                                        ,Renal_Failure
                                        ,Death
                                        ,Stroke
                                        ,RFDS
                                        ,RFD
                                        ,DS
                                        ,RFS

                                        
                           ))

```

### Removing columns where 100% is NA

```{r}
#abp_mean_all_data <- abp_mean_all_data[,colSums(is.na(abp_mean_all_data))<nrow(abp_mean_all_data)]
```


### Imputing mising values in mean_ef_median_imp

there were some mising values in mean_ef_median_imp, imputing mean_ef_median_imp for 81 cases missing

```{r}
abp_mean_all_data$mean_ef_median_imp[is.na(abp_mean_all_data$mean_ef_median_imp)]<-round(mean(abp_mean_all_data$mean_ef_median_imp,na.rm = T),2)
```

### Maping variables

We don't need all the variables in our dataset to be numbers anymore.

```{r}
#abp_mean_all_data$Gender<-plyr::revalue(abp_mean_all_data$Gender,c('Female'=1,'Male'=2))
#abp_mean_all_data$Category<-plyr::revalue(abp_mean_all_data$Category,c('Aortic Surgery'=1,'CABG'=2,'CABG + Valve'=3,'Other'=3,'Valve'=4))
```

### Generating report

```{r}
view(dfSummary(abp_mean_all_data))
```

### Writing dataframe

```{r}
write.csv(abp_mean_all_data,'abp_mean_all_data.csv',row.names = F)
```
 

