---
title: "01_data_exploration"
author: "Miguel Armengol"
date: "2/20/2019"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(feather)
library(knitr)
library(plotly)
library(sqldf)
library(zoo)
library(data.table)
library(summarytools)
library(imputeTS)
library(pracma)
library(entropy)
```

# Importing CPB times data 

Data was imported from the project "Nonlinear Association of Intraoperative Vasopressor-Inotrope Dose and Hypotension Duration with Adverse Outcomes following Cardiac Surgery" where these variables were generated, code on how these variables were generated is available on the project [repo](https://github.com/theonesp/vasopressor_dose_mae/).

- From our Vitalsign table, there are 390 Cases with no CPB Times Info.
- Meaning there are 5318 Cases with all the info we need
- We are not interested on Cases with NO CPB informations, thus we exclude this cases.

*TIMEZONES NEED TO BE THE SAME ACROSS DATASETS AND COLUMNS*

```{r}
AllPeriodsDF <- readRDS("~/MEGA/Boston/BIDMC-Harvard/bp_nn/AllPeriodsDF.rds")
# before working with the dataset we need to make sure we are using the same timezone across all our columns and datasets
sapply(AllPeriodsDF, function(x) attr(x,"tzone"))
desiredtimezone<-attr(abp_mean_expanded$Time,"tzone")

AllPeriodsDF$CPB_Begin_Time<-as.character(AllPeriodsDF$CPB_Begin_Time)
AllPeriodsDF$CPB_Begin_Time<-as.POSIXct(AllPeriodsDF$CPB_Begin_Time, format='%Y-%m-%d %H:%M:%S', tz=desiredtimezone)

AllPeriodsDF$CPB_End_Time<-as.character(AllPeriodsDF$CPB_End_Time)
AllPeriodsDF$CPB_End_Time<-as.POSIXct(AllPeriodsDF$CPB_End_Time, format='%Y-%m-%d %H:%M:%S', tz=desiredtimezone)
```

ingest vitasigns_df!!!!!!! the other timestamps are corrupted

# abp_mean_cpb

*We are going to include only those cases we defined match our inclusion criteria as explained on '02_dataset_creation_full'*

We use abp_mean_expanded as created in 02_dataset_creation_full.Rmd so values have already been processed:
 - Excluding gaps > 5 minutes or surgery length < 60 minutes.
 - Expanding data points when several missing time-stamps are found.
 - Artifact Removal.
 - Imputing intra-surgery missing data: Median imputation.

# Subseting the data only to include CPB time

```{r}
abp_mean_expanded_cpb<-inner_join(  abp_mean_expanded,  AllPeriodsDF)%>%
  filter(Time>=CPB_Begin_Time & Time<=CPB_End_Time)%>%
  select(Time, Case_Name, Value,Value_imputed)
```

### Reshaping

```{r}
# I'm creating a counter column so we know how many datapoints we have per case (we'll later use this for reshaping the dataframe)
abp_mean_expanded_cpb_count<-abp_mean_expanded_cpb %>%
  group_by(Case_Name) %>%
  mutate(
    ones = 1,
    counter = cumsum(ones)
    )

# Reshaping the table so we have 1 row per patient and one column each data point
abp_mean_reshaped_cpb<-dcast(abp_mean_expanded_cpb_count, Case_Name ~ counter, value.var = "Value_imputed")
```

### Imputing data with a special value.

LVP and linear interpolation methods can also be tried for imputation.

Instead of removing the rows with NaN values, we can replace all NaN values with a specific value that does not appear naturally in the input, such as -1.
The problem can be modeled as-is and we can encourage the model to learn that a specific value means “missing.” 

Ref: https://machinelearningmastery.com/handle-missing-timesteps-sequence-prediction-problems-python/

```{r}
abp_mean_reshaped_cpb[is.na(abp_mean_reshaped_cpb)]<-(-1)
```

### Entropy

```{r}
abp_mean_reshaped_cpb['entropy']<-round(apply(abp_mean_reshaped_cpb[2:ncol(abp_mean_reshaped_cpb)], 1 , entropy ) , 2)
```

Each observation (row) corresponds to one case, there are 3650 data points (variables=columns), 1 every 15 seconds. 
This is because the longest case contains 3650 data points using ST 15 seconds. 
Cases shorter than this one will contain NA from the last time-stamp of the surgery until data-point number 3650.

### Joining with other data-sets
 
```{r}
abp_mean_cpb_all_data<-merge(abp_mean_reshaped_cpb,
                         FinalHemodynamicsDataset%>%
                          dplyr::select( Case_Name
                                        ,Gender
                                        ,Age
                                        ,Category
                                        
                                        ,duration_Surgery_min
                                        ,duration_CPB_min
                                        ,duration_outsideCPB_min
                                        
                                        ,STS_RiskAlgorithm
                                        ,STS_RiskAlgorithm_multiple_imp
                                        ,STS_RiskAlgorithm_median_imp
                                        ,total_pressors_formula_mg
                                        
                                        # categorical variables
                                        ,Q_total_pressors_formula_mg
                                        ,T_STS_RiskAlgorithm_median_imp
                                        ,T_STS_RiskAlgorithm_multiple_imp
                                        ,Q_AUC65MAP_total_mmHgmin
                                        ,Q_TWA65MAP_total_mmHg
                                        ,Q_AUC65MAP_CPB_mmHgmin
                                        ,Q_AUC65MAP_outsideCPB_mmHgmin
                                        ,Q_TWA65MAP_CPB_mmHg
                                        ,Q_TWA65MAP_outsideCPB_mmHg
                                        
                                                                                
                                        # TWA variables
                                        ,TWA65MAP_total_mmHg
                                        ,TWA65MAP_CPB_mmHg
                                        ,TWA65MAP_outsideCPB_mmHg
                                        

                                                                                                                                                                             ,delta_HCT_pct_median_imp
                                        ,mean_ef_median_imp
                                        ,Cross_Clamp_Time_imp
                                        
                                        ,AUC65MAP_total_mmHgmin
                                        ,AUC65MAP_CPB_mmHgmin
                                        ,AUC65MAP_outsideCPB_mmHgmin
                                        
                                        # Outcomes
                                        ,Any_MAE
                                        ,Renal_Failure
                                        ,Death
                                        ,Stroke
                                        ,RFDS
                                        ,RFD
                                        ,DS
                                        ,RFS

                                        
                           ))
```

### Imputing mising values in mean_ef_median_imp

there were some mising values in mean_ef_median_imp, imputing mean_ef_median_imp for 81 cases missing

```{r}
abp_mean_cpb_all_data$mean_ef_median_imp[is.na(abp_mean_cpb_all_data$mean_ef_median_imp)]<-round(mean(abp_mean_cpb_all_data$mean_ef_median_imp,na.rm = T),2)
```

### Generating report

```{r eval=FALSE, include=FALSE}
view(dfSummary(abp_mean_cpb_all_data))
```

### Exporting df

```{r eval=FALSE, include=FALSE}
write.csv(abp_mean_cpb_all_data,'abp_mean_cpb_all_data.csv',row.names = F)
``` 

